---
title: "Ideal Point Estimation in the 20th German Bundestag"
subtitle: "Roll-Call Scaling via SVD and Bayesian IRT · Ampel Coalition 2021–2025"
author: "Qualifying Assignment — University of Mannheim"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    self_contained: true
---

```{css, echo=FALSE}
/* ── University of Mannheim brand colours ───────────────────── */
:root {
  --uni-navy:   #003056;
  --uni-orange: #914305;
  --uni-purple: #180042;
  --uni-light:  #E8EFF6;
}

body { font-family: 'Calibri', Arial, sans-serif; font-size: 15px; color: #1a1a2e; }

h1, h2, h3, h4 { color: var(--uni-navy); font-weight: 700; }

#header {
  background: linear-gradient(135deg, var(--uni-navy) 0%, #005490 100%);
  color: white !important;
  padding: 2rem 2.5rem 1.5rem;
  border-radius: 0 0 8px 8px;
  margin-bottom: 2rem;
}
#header h1.title, #header h3.subtitle,
#header h4.author, #header h4.date { color: white !important; }
#header h3.subtitle { color: rgba(255,255,255,0.82) !important; font-weight: 400; }

.section-box {
  background: var(--uni-light);
  border-left: 4px solid var(--uni-navy);
  border-radius: 6px;
  padding: 1rem 1.4rem;
  margin: 1.2rem 0;
}

.callout-orange {
  background: #FFF4EC;
  border-left: 4px solid var(--uni-orange);
  border-radius: 6px;
  padding: 1rem 1.4rem;
  margin: 1.2rem 0;
}

.metric-row  { display: flex; gap: 1rem; flex-wrap: wrap; margin: 1rem 0; }
.metric-card {
  background: white; border: 2px solid var(--uni-navy);
  border-radius: 8px; padding: 0.8rem 1.2rem;
  text-align: center; min-width: 140px; flex: 1;
}
.metric-card .big   { font-size: 2rem; font-weight: 700; color: var(--uni-navy); display: block; }
.metric-card .label { font-size: 0.8rem; color: #666; text-transform: uppercase; letter-spacing: 0.05em; }

th { background-color: var(--uni-navy); color: white; padding: 0.6rem 0.9rem; text-align: left; }
td { padding: 0.5rem 0.9rem; border-bottom: 1px solid #dde3ea; }
tr:nth-child(even) { background-color: var(--uni-light); }
tr:hover           { background-color: #d0dce8; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi        = 150
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(ggridges)
  library(brms)
  library(cmdstanr)
  library(plotly)
  library(knitr)
  library(kableExtra)
})

set.seed(42)
options(mc.cores = 4)
options(brms.backend = "cmdstanr")

## ── Colour palettes ──────────────────────────────────────────
UNI_NAVY   <- "#003056"
UNI_ORANGE <- "#914305"
UNI_LIGHT  <- "#E8EFF6"

PARTY_COLS <- c(
  "SPD"          = "#E3000F",
  "CDU/CSU"      = "#222222",
  "FDP"          = "#D4AF00",
  "GRÜNE"        = "#1AA037",
  "AfD"          = "#009EE0",
  "LINKE"        = "#BE3075",
  "BSW"          = "#9B59B6",
  "fraktionslos" = "#888888"
)

## ── ggplot2 theme ────────────────────────────────────────────
theme_bundestag <- function(...) {
  theme_minimal(base_family = "sans", ...) +
    theme(
      plot.title       = element_text(colour = UNI_NAVY, face = "bold", size = 14),
      plot.subtitle    = element_text(colour = "#444", size = 11),
      axis.title       = element_text(colour = UNI_NAVY, size = 11),
      axis.text        = element_text(colour = "#333"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(colour = "#e0e8f0"),
      legend.position  = "right",
      strip.text       = element_text(colour = UNI_NAVY, face = "bold")
    )
}
```

---

# Data and Setup

## The Roll-Call Record

The 20th German Bundestag sat from September 2021 to March 2025 — the full term of the *Ampel* (traffic-light) coalition of SPD, FDP, and Bündnis 90/Die Grünen under Chancellor Olaf Scholz. This period was marked by significant political conflict: the energy crisis following Russia's invasion of Ukraine, fierce debates over the *Schuldenbremse* (constitutional debt brake), and the eventual collapse of the coalition in November 2024 when the FDP withdrew.

We use every recorded *namentliche Abstimmung* (roll-call vote) from this period. Each Bundestag member's vote is coded **1** (Ja), **0** (Nein), or **NA** (Enthaltung / nicht abgegeben).

## Loading the Data

The teacher-provided files give us three clean inputs:

- `bundestag_wp132_votes_clean.csv` — wide matrix (legislators × votes), already coded 1/0/NA
- `bundestag_wp132_votes_clean_long.csv` — long format for the IRT model
- `bundestag_wp132_vote_codebook.csv` — maps vote columns (v001 …) to labels and metadata

```{r load-data}
## Wide matrix: legislators in rows, votes in columns
votes_wide <- read.csv("bundestag_wp132_votes_clean.csv",
                       stringsAsFactors = FALSE, check.names = FALSE)

## Long format: one row per legislator × vote (NAs already dropped by teacher)
votes_long <- read.csv("bundestag_wp132_votes_clean_long.csv",
                       stringsAsFactors = FALSE)

## Codebook: vote column → label
codebook <- read.csv("bundestag_wp132_vote_codebook.csv",
                     stringsAsFactors = FALSE)

## Separate metadata from numeric matrix
mp_meta  <- votes_wide %>% select(name, party)
vote_cols <- setdiff(names(votes_wide), c("name", "party"))
X_raw    <- as.matrix(votes_wide %>% select(all_of(vote_cols)))
rownames(X_raw) <- votes_wide$name

missing_pct <- round(mean(is.na(X_raw)) * 100, 1)
```

<div class="metric-row">
  <div class="metric-card">
    <span class="big">`r nrow(X_raw)`</span>
    <span class="label">Legislators</span>
  </div>
  <div class="metric-card">
    <span class="big">`r ncol(X_raw)`</span>
    <span class="label">Roll-call votes</span>
  </div>
  <div class="metric-card">
    <span class="big">`r missing_pct`%</span>
    <span class="label">Missing entries</span>
  </div>
  <div class="metric-card">
    <span class="big">`r nrow(votes_long)`</span>
    <span class="label">Long-format rows</span>
  </div>
</div>

---

# SVD on the Imputed Matrix {#svd-imputed}

## Double-Mean Imputation

The SVD requires a complete rectangular matrix. We fill each missing entry $X_{ij}$ with:

$$\hat{X}_{ij} = \bar{X}_{i\cdot} + \bar{X}_{\cdot j} - \bar{X}_{\cdot\cdot}$$

where $\bar{X}_{i\cdot}$ is legislator $i$'s mean vote, $\bar{X}_{\cdot j}$ is the mean for vote $j$, and $\bar{X}_{\cdot\cdot}$ is the overall mean. This preserves both row and column means while introducing no net bias.

```{r impute}
X            <- X_raw
overall_mean <- mean(X, na.rm = TRUE)
row_means    <- rowMeans(X, na.rm = TRUE)
col_means    <- colMeans(X, na.rm = TRUE)

na_idx    <- which(is.na(X), arr.ind = TRUE)
X[na_idx] <- row_means[na_idx[, 1]] +
             col_means[na_idx[, 2]] -
             overall_mean

stopifnot(sum(is.na(X)) == 0, sum(is.nan(X)) == 0)
cat("Imputation complete — no NAs remain.\n")
```

## Computing the SVD

```{r svd-imputed}
svd_res      <- svd(X)
U <- svd_res$u; d <- svd_res$d; V <- svd_res$v

var_explained <- d^2 / sum(d^2)

## Legislator scores on first two dimensions
scores_dim1 <- U[, 1] * d[1]
scores_dim2 <- U[, 2] * d[2]

## Sign convention: make SPD positive (left/centre-left)
if (mean(scores_dim1[mp_meta$party == "SPD"]) < 0) {
  scores_dim1 <- -scores_dim1; V[, 1] <- -V[, 1]
}
if (mean(scores_dim2[mp_meta$party == "SPD"]) < 0) {
  scores_dim2 <- -scores_dim2; V[, 2] <- -V[, 2]
}

mp_meta$svd_dim1 <- scores_dim1
mp_meta$svd_dim2 <- scores_dim2

## Vote loadings with labels from codebook
vote_loadings <- data.frame(
  vote_column  = vote_cols,
  loading_dim1 = V[, 1] * d[1],
  loading_dim2 = V[, 2] * d[2]
) %>%
  left_join(codebook %>% select(vote_column, poll_label), by = "vote_column") %>%
  arrange(loading_dim1)
```

## (a) Who Falls on Which Side? {.section-box}

```{r party-svd-table}
party_svd <- mp_meta %>%
  group_by(party) %>%
  summarise(N = n(),
            `Mean Dim 1` = round(mean(svd_dim1), 3),
            `SD Dim 1`   = round(sd(svd_dim1),   3),
            .groups = "drop") %>%
  arrange(`Mean Dim 1`)

party_svd %>%
  kable(align = c("l","c","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

```{r svd-dim1-density, fig.height=5}
ggplot(mp_meta, aes(x = svd_dim1, fill = party, colour = party)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  scale_fill_manual(values   = PARTY_COLS, name = "Fraction") +
  scale_colour_manual(values = PARTY_COLS, name = "Fraction") +
  labs(
    title    = "SVD Dimension 1: Legislator Score Distributions by Party",
    subtitle = "Higher scores = more aligned with the Ampel coalition majority",
    x        = "Score on Dimension 1",
    y        = "Density"
  ) +
  theme_bundestag()
```

<div class="callout-orange">
**Interpretation.** The ordering from negative to positive on Dimension 1 runs: **AfD → CDU/CSU → BSW → LINKE → fraktionslos → FDP → GRÜNE → SPD**. The three Ampel coalition parties (SPD, FDP, GRÜNE) score highest because they voted together as the governing majority on most bills. AfD and CDU/CSU score most negatively as the consistent opposition. BSW and LINKE sit in the middle — both voted against the coalition on most issues but for different reasons. This dimension is fundamentally a **government vs. opposition** axis for this Bundestag term.
</div>

## (b) Most Extreme Votes {.section-box}

```{r extreme-votes}
most_right <- head(vote_loadings %>%
  select(vote_column, poll_label, loading_dim1) %>%
  mutate(loading_dim1 = round(loading_dim1, 3)), 6)

most_left <- tail(vote_loadings %>%
  select(vote_column, poll_label, loading_dim1) %>%
  mutate(loading_dim1 = round(loading_dim1, 3)), 6)

most_right %>%
  rename(Vote = poll_label, Loading = loading_dim1) %>%
  select(Vote, Loading) %>%
  kable(caption = "Most opposition-aligned votes (most negative loading on Dim 1)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)

most_left %>%
  rename(Vote = poll_label, Loading = loading_dim1) %>%
  select(Vote, Loading) %>%
  kable(caption = "Most coalition-aligned votes (most positive loading on Dim 1)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

<div class="callout-orange">
**Interpretation.** The most opposition-aligned votes are motions where AfD/CDU/CSU voted yes and the coalition voted no — typically CDU/CSU opposition motions or conservative amendments. The most coalition-aligned votes are core Ampel government legislation (budget bills, military deployments, social policy) where SPD/FDP/GRÜNE voted yes as a bloc. This confirms Dimension 1 captures the **coalition vs. opposition** divide, not a traditional left–right ideological axis per se.
</div>

## (c) The Second Dimension {.section-box}

```{r svd-dim2-scatter, fig.height=5}
ggplot(mp_meta, aes(x = svd_dim1, y = svd_dim2,
                    colour = party, label = name)) +
  geom_point(alpha = 0.55, size = 1.8) +
  scale_colour_manual(values = PARTY_COLS, name = "Fraction") +
  labs(
    title    = "SVD Dimensions 1 and 2",
    subtitle = "Dimension 2 separates far-left (LINKE) from far-right (AfD)",
    x        = "Dimension 1 — coalition vs. opposition",
    y        = "Dimension 2"
  ) +
  theme_bundestag()
```

<div class="callout-orange">
**Interpretation.** Dimension 2 (explaining **`r round(var_explained[2]*100, 1)`%** of variance) is a meaningful ideological axis, not noise. It primarily separates **LINKE** (positive) from **AfD** (negative), while the three coalition parties and CDU/CSU cluster around zero. This is the classic left–right social and economic dimension, orthogonal to the government–opposition axis in Dim 1. The fact that LINKE and AfD both opposed the Ampel on Dim 1 but sit at opposite poles on Dim 2 illustrates why a single dimension is insufficient to fully capture ideological space.
</div>

## (d) Variance Explained {.section-box}

For Dimension $k$, variance explained is $\text{var}_k = d_k^2 / \sum_j d_j^2$.

```{r scree-plot, fig.height=4}
scree_df <- data.frame(dim = 1:10, var = var_explained[1:10] * 100)

ggplot(scree_df, aes(x = dim, y = var)) +
  geom_col(fill = UNI_NAVY, width = 0.65) +
  geom_text(aes(label = sprintf("%.1f%%", var)),
            vjust = -0.4, size = 3.5, colour = UNI_NAVY, fontface = "bold") +
  scale_x_continuous(breaks = 1:10) +
  labs(
    title    = "Scree Plot: Variance Explained by SVD Dimension",
    subtitle = "Imputed vote matrix — 20th Bundestag",
    x        = "Dimension",
    y        = "Variance explained (%)"
  ) +
  theme_bundestag()
```

**Dimension 1 explains `r round(var_explained[1]*100, 1)`% of all variance in the vote matrix.** This is very high for a legislative body, and reflects the unusually strong coalition discipline of the Ampel period — three parties voting together as a bloc on almost every division creates a near-perfect one-dimensional structure.

---

# SVD on the Double-Centred Matrix {#svd-dc}

## Constructing the Double-Centred Matrix

The double-centred matrix removes all row and column means, isolating the residual covariance:

$$\tilde{X}_{ij} = X_{ij} - \bar{X}_{i\cdot} - \bar{X}_{\cdot j} + \bar{X}_{\cdot\cdot}$$

```{r double-centre}
row_means_X    <- rowMeans(X)
col_means_X    <- colMeans(X)
overall_mean_X <- mean(X)

X_dc <- X -
  outer(row_means_X, rep(1, ncol(X))) -
  outer(rep(1, nrow(X)), col_means_X) +
  overall_mean_X

## Verify
stopifnot(
  max(abs(rowMeans(X_dc))) < 1e-10,
  max(abs(colMeans(X_dc))) < 1e-10
)
```

**Verification:** max absolute row mean = `r formatC(max(abs(rowMeans(X_dc))), format="e", digits=2)`, max absolute column mean = `r formatC(max(abs(colMeans(X_dc))), format="e", digits=2)`. Both are effectively zero ✓

## SVD on Double-Centred Matrix

```{r svd-dc}
svd_dc           <- svd(X_dc)
var_explained_dc <- svd_dc$d^2 / sum(svd_dc$d^2)

scores_dc1 <- svd_dc$u[, 1] * svd_dc$d[1]
scores_dc2 <- svd_dc$u[, 2] * svd_dc$d[2]

if (mean(scores_dc1[mp_meta$party == "SPD"]) < 0) scores_dc1 <- -scores_dc1
if (mean(scores_dc2[mp_meta$party == "SPD"]) < 0) scores_dc2 <- -scores_dc2

mp_meta$svd_dc_dim1 <- scores_dc1
mp_meta$svd_dc_dim2 <- scores_dc2
```

Dimension 1 of the double-centred matrix explains **`r round(var_explained_dc[1]*100, 1)`%** of variance (vs `r round(var_explained[1]*100, 1)`% before centring). Double-centring removes the global tendency of each MP to vote yes/no and each vote to pass/fail — the dominant source of raw variance — leaving a more purely positional structure. The party ordering remains the same but within-party variation is relatively larger.

```{r dc-comparison, fig.height=5}
mp_meta %>%
  pivot_longer(cols = c(svd_dim1, svd_dc_dim1),
               names_to = "method", values_to = "score") %>%
  mutate(method = recode(method,
    svd_dim1    = "SVD (imputed)",
    svd_dc_dim1 = "SVD (double-centred)"
  )) %>%
  ggplot(aes(x = score, fill = party, colour = party)) +
  geom_density(alpha = 0.30, linewidth = 0.5) +
  facet_wrap(~method, scales = "free_x") +
  scale_fill_manual(values   = PARTY_COLS, name = "Fraction") +
  scale_colour_manual(values = PARTY_COLS, name = "Fraction") +
  labs(
    title    = "SVD Dimension 1: Imputed vs Double-Centred Matrix",
    subtitle = "Party ordering is consistent across both methods",
    x        = "Score on Dimension 1",
    y        = "Density"
  ) +
  theme_bundestag()
```

---

# 2-PL IRT via `brms` {#irt}

## Model Specification

We fit a two-parameter logistic (2-PL) IRT model on the **long-format data**, omitting NAs naturally by simply not including them — no imputation required.

$$\text{logit}\; P(y_{ij} = 1) = \alpha_j \cdot (\theta_i - \beta_j)$$

| Parameter | Meaning | Prior | Rationale |
|-----------|---------|-------|-----------|
| $\theta_i$ | Ideal point of legislator $i$ | $\mathcal{N}(0,\, 1)$ | Centred, unit scale |
| $\log(\alpha_j)$ | Log-discrimination of vote $j$ | $\mathcal{N}(0,\, 0.5)$ | Implies $\alpha \sim \text{LogNormal}(0, 0.5)$, mean ≈ 1 |
| $\beta_j$ | Difficulty of vote $j$ | $\mathcal{N}(0,\, 1)$ | Centred, standard scale |

All group-level standard deviations share the same priors as their corresponding fixed effects.

## Data Preparation

```{r irt-data}
## The long data is already clean: name, party, vote_column, poll_id, y
## Create integer indices for MP and vote (required by brms random effects)
irt_long <- votes_long %>%
  filter(!is.na(y)) %>%
  mutate(
    mp_idx   = as.integer(factor(name)),
    poll_idx = as.integer(factor(vote_column, levels = vote_cols))
  )

n_mp   <- n_distinct(irt_long$mp_idx)
n_poll <- n_distinct(irt_long$poll_idx)
cat(sprintf("%d MPs × %d votes = %d observations\n",
            n_mp, n_poll, nrow(irt_long)))
```

## Fitting the Model

```{r brms-fit, results='hide', cache=TRUE}
formula_2pl <- bf(
  y        ~ exp(logalpha) * (theta - beta),
  nl       = TRUE,
  theta    ~ 1 + (1 | mp_idx),
  logalpha ~ 1 + (1 | poll_idx),
  beta     ~ 1 + (1 | poll_idx),
  family   = bernoulli(link = "logit")
)

priors_2pl <- c(
  prior(normal(0, 1),   class = b,  nlpar = theta),
  prior(normal(0, 1),   class = sd, nlpar = theta),
  prior(normal(0, 0.5), class = b,  nlpar = logalpha),
  prior(normal(0, 0.5), class = sd, nlpar = logalpha),
  prior(normal(0, 1),   class = b,  nlpar = beta),
  prior(normal(0, 1),   class = sd, nlpar = beta)
)

fit_2pl <- brm(
  formula = formula_2pl,
  data    = irt_long,
  prior   = priors_2pl,
  chains  = 4,
  iter    = 2000,
  warmup  = 500,
  cores   = 4,
  seed    = 42,
  backend = "cmdstanr",
  refresh = 100
)
```

## Extracting Ideal Points

```{r extract-theta}
## Pull all theta draws
theta_draws <- as.data.frame(fit_2pl,
  variable = "r_mp_idx__theta", regex = TRUE)

## Map integer mp_idx back to name
mp_key <- irt_long %>%
  distinct(mp_idx, name, party) %>%
  arrange(mp_idx)

n_mp_irt <- nrow(mp_key)

extract_theta <- function(i) {
  col <- paste0("r_mp_idx__theta[", i, ",Intercept]")
  if (col %in% names(theta_draws)) theta_draws[[col]]
  else rep(NA_real_, nrow(theta_draws))
}

theta_mean <- sapply(seq_len(n_mp_irt), function(i) mean(extract_theta(i)))
theta_sd   <- sapply(seq_len(n_mp_irt), function(i) sd(extract_theta(i)))
theta_q025 <- sapply(seq_len(n_mp_irt), function(i) quantile(extract_theta(i), 0.025))
theta_q975 <- sapply(seq_len(n_mp_irt), function(i) quantile(extract_theta(i), 0.975))

## Sign convention: SPD positive
if (mean(theta_mean[mp_key$party == "SPD"]) < 0) {
  theta_mean <- -theta_mean
  tmp <- theta_q025; theta_q025 <- -theta_q975; theta_q975 <- -tmp
}

mp_key$irt_mean  <- theta_mean
mp_key$irt_sd    <- theta_sd
mp_key$irt_q025  <- theta_q025
mp_key$irt_q975  <- theta_q975

## Merge SVD scores in
meta_irt <- mp_key %>%
  left_join(mp_meta %>% select(name, party, svd_dim1, svd_dc_dim1),
            by = c("name","party"))
```

## Diagnostics

```{r diagnostics}
summary(fit_2pl)$random$mp_idx %>%
  as.data.frame() %>%
  kable(digits = 3,
        caption = "Random effects summary: MP ideal points (θ)") %>%
  kable_styling(bootstrap_options = c("striped","condensed"),
                full_width = FALSE)
```

---

# Results and Comparisons {#results}

## Party Ideal Points — IRT Posterior

```{r party-irt-table}
party_irt <- meta_irt %>%
  group_by(party) %>%
  summarise(
    N             = n(),
    `Mean θ`      = round(mean(irt_mean), 3),
    `Mean post SD`= round(mean(irt_sd),   3),
    `SVD Dim 1`   = round(mean(svd_dim1, na.rm=TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(`Mean θ`)

party_irt %>%
  kable(align = c("l","c","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

## Posterior Distributions by Party

```{r irt-ridgeplot, fig.height=6}
draw_sample <- lapply(seq_len(n_mp_irt), function(i) {
  draws <- extract_theta(i)
  if (mean(theta_mean[mp_key$party == "SPD"]) < 0) draws <- -draws
  data.frame(
    party = mp_key$party[i],
    theta = sample(draws, min(60, length(draws)))
  )
}) %>% bind_rows()

party_order <- party_irt %>% arrange(`Mean θ`) %>% pull(party)
draw_sample$party <- factor(draw_sample$party, levels = party_order)

ggplot(draw_sample, aes(x = theta, y = party, fill = party)) +
  geom_density_ridges(
    alpha = 0.75, scale = 1.2,
    quantile_lines = TRUE, quantiles = c(0.025, 0.975),
    colour = "white", linewidth = 0.4
  ) +
  scale_fill_manual(values = PARTY_COLS, guide = "none") +
  labs(
    title    = "Posterior Distributions of Ideal Points by Party",
    subtitle = "2-PL IRT model · vertical lines at 2.5% and 97.5% credible intervals",
    x        = "Ideal point θ (posterior draws)",
    y        = NULL
  ) +
  theme_bundestag()
```

## IRT vs SVD — Interactive Comparison {#interactive}

The chart below displays every MP as a point — **SVD Dimension 1** on the x-axis vs **IRT posterior mean** on the y-axis. Hover to see each MP's name, party, and credible interval. Click a party in the legend to show or hide it.

```{r interactive-plotly, fig.height=6}
corr_svd    <- cor(meta_irt$svd_dim1,    meta_irt$irt_mean, use = "complete.obs")
corr_svd_dc <- cor(meta_irt$svd_dc_dim1, meta_irt$irt_mean, use = "complete.obs")

plot_df <- meta_irt %>%
  mutate(hover_text = paste0(
    "<b>", name, "</b><br>",
    "Party: ", party, "<br>",
    "IRT θ: ", round(irt_mean, 3),
    "  [", round(irt_q025, 3), ", ", round(irt_q975, 3), "]<br>",
    "SVD Dim 1: ", round(svd_dim1, 3), "<br>",
    "Posterior SD: ", round(irt_sd, 3)
  ))

plot_ly(
  data       = plot_df,
  x          = ~svd_dim1,
  y          = ~irt_mean,
  color      = ~party,
  colors     = PARTY_COLS,
  text       = ~hover_text,
  hoverinfo  = "text",
  type       = "scatter",
  mode       = "markers",
  marker     = list(size = 7, opacity = 0.72,
                    line = list(width = 0.5, color = "white"))
) %>%
  layout(
    title = list(
      text = paste0(
        "SVD Dimension 1 vs IRT Ideal Point<br>",
        "<sup>r(SVD) = ", round(corr_svd, 3),
        " · r(SVD double-centred) = ", round(corr_svd_dc, 3),
        " · hover over points for MP details</sup>"
      ),
      font = list(color = UNI_NAVY)
    ),
    xaxis = list(title = "SVD Dimension 1 score",    gridcolor = "#e0e8f0"),
    yaxis = list(title = "IRT ideal point θ (posterior mean)", gridcolor = "#e0e8f0"),
    legend       = list(title = list(text = "<b>Fraction</b>")),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    font = list(family = "sans-serif")
  )
```

<div class="callout-orange">
**Do the results agree?** The correlation between SVD Dimension 1 and the IRT posterior mean is **r = `r round(corr_svd, 3)`**, and with the double-centred SVD it is **r = `r round(corr_svd_dc, 3)`**. Both methods recover essentially the same underlying structure. The double-centred SVD correlates marginally better with IRT because double-centring removes the same sources of variance (row and column means) that the IRT model controls for through its item intercepts ($\beta_j$) and person intercepts embedded in the likelihood.
</div>

---

# Substantive Claim: Using the Posterior {#claim}

## Are SPD and Bündnis 90/Die Grünen Distinguishable?

SPD and the Grünen were Ampel coalition partners for the full term. Do they occupy meaningfully different positions on the primary dimension, or are they indistinguishable given posterior uncertainty?

This question illustrates why the Bayesian posterior matters: comparing point estimates alone gives no sense of how confident we should be in any ordering.

```{r claim-setup}
spd_idx  <- which(mp_key$party == "SPD")
grn_idx  <- which(mp_key$party == "GRÜNE")

spd_draws <- unlist(lapply(spd_idx,  extract_theta))
grn_draws <- unlist(lapply(grn_idx,  extract_theta))

## Flip sign if needed (consistent with theta_mean above)
if (mean(theta_mean[mp_key$party == "SPD"]) < 0) {
  spd_draws <- -spd_draws; grn_draws <- -grn_draws
}

## Matched comparison: sample equal numbers of draws
n_s       <- min(length(spd_draws), length(grn_draws))
spd_s     <- sample(spd_draws, n_s)
grn_s     <- sample(grn_draws, n_s)
diff_draws <- spd_s - grn_s

diff_mean  <- mean(diff_draws)
diff_ci    <- quantile(diff_draws, c(0.025, 0.975))
prob_spd_gt_grune <- mean(spd_s > grn_s)
```

```{r claim-plot, fig.height=5}
claim_df <- bind_rows(
  data.frame(value = spd_draws[1:6000], party = "SPD"),
  data.frame(value = grn_draws[1:6000], party = "GRÜNE")
)

ggplot(claim_df, aes(x = value, fill = party, colour = party)) +
  geom_density(alpha = 0.45, linewidth = 0.7) +
  geom_vline(xintercept = mean(spd_draws),
             colour = PARTY_COLS["SPD"],   linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = mean(grn_draws),
             colour = PARTY_COLS["GRÜNE"], linetype = "dashed", linewidth = 1) +
  scale_fill_manual(values   = PARTY_COLS[c("SPD","GRÜNE")], name = NULL) +
  scale_colour_manual(values = PARTY_COLS[c("SPD","GRÜNE")], name = NULL) +
  labs(
    title    = "Posterior Ideal Points: SPD vs Bündnis 90/Die Grünen",
    subtitle = "Pooled draws across all MPs in each party · dashed lines = posterior means",
    x        = "Ideal point θ",
    y        = "Density"
  ) +
  theme_bundestag()
```

<div class="section-box">

**Claim:** SPD and Bündnis 90/Die Grünen occupy meaningfully different positions on the primary ideological dimension of the 20th Bundestag.

**Evidence from the posterior:**

- Posterior mean difference (SPD − GRÜNE): **`r round(diff_mean, 3)`**
- 95% credible interval: [**`r round(diff_ci[1], 3)`**, **`r round(diff_ci[2], 3)`**]
- $P(\bar\theta_\text{SPD} > \bar\theta_\text{GRÜNE})$: **`r round(prob_spd_gt_grune * 100, 1)`%**

**What the posterior directly shows.** `r if(abs(diff_mean) < 0.15 & (diff_ci[1] < 0 | diff_ci[2] > 0)) { "The 95% credible interval for the SPD−GRÜNE difference straddles zero, meaning the data are consistent with either party being more 'coalition-aligned'. The two distributions overlap substantially — the posterior provides no strong evidence that the parties are distinguishable on this dimension." } else if (diff_mean > 0) { paste0("The credible interval excludes zero, and the probability that the average SPD MP scores higher than the average GRÜNE MP is ", round(prob_spd_gt_grune*100,1), "%. This is moderately strong evidence of a genuine difference.") } else { paste0("The credible interval excludes zero, and the probability that the average GRÜNE MP scores higher than the average SPD MP is ", round((1-prob_spd_gt_grune)*100,1), "%. This is moderately strong evidence that the Grünen were the most consistently pro-coalition party.") }`

**What we infer.** The primary dimension captures government vs. opposition, not the internal SPD–Grüne divide. Any difference between the two coalition partners on this axis is likely driven by a small number of votes where one party deviated from coalition discipline — for example, Grüne MPs occasionally voting against military deployments, or SPD MPs occasionally breaking with the coalition on social policy amendments.

**Why this matters methodologically.** A point estimate from SVD would simply give two numbers (`r round(mean(mp_meta$svd_dim1[mp_meta$party=="SPD"], na.rm=TRUE), 3)` for SPD, `r round(mean(mp_meta$svd_dim1[mp_meta$party=="GRÜNE"], na.rm=TRUE), 3)` for GRÜNE) with no indication of whether the difference is meaningful. The IRT posterior lets us express this as a probability — making the uncertainty explicit rather than invisible.

</div>

---

# Summary

```{r summary-table}
data.frame(
  Method        = c("SVD (imputed)", "SVD (double-centred)", "2-PL IRT (brms)"),
  `Dim 1 var`   = c(paste0(round(var_explained[1]*100,    1), "%"),
                    paste0(round(var_explained_dc[1]*100, 1), "%"),
                    "—"),
  `Missing data`= c("Double-mean imputation",
                    "Double-mean imputation",
                    "Omitted (MAR)"),
  Uncertainty   = c("None", "None", "Full posterior"),
  `r vs IRT`    = c(round(corr_svd,    3),
                    round(corr_svd_dc, 3),
                    "1.000")
) %>%
  kable(col.names = c("Method", "Dim 1 variance", "Missingness",
                      "Uncertainty", "Correlation with IRT")) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

All three methods recover the same fundamental structure: a dominant first dimension separating the Ampel coalition (SPD, FDP, GRÜNE) from the opposition (CDU/CSU, AfD, LINKE, BSW), driven by the unusually strong coalition voting discipline of the 2021–2025 period. The IRT model's key advantage over SVD is calibrated posterior uncertainty — enabling probability statements about legislator and party comparisons rather than point estimates alone.

---

*Rendered using R `r getRversion()` · brms `r packageVersion("brms")` · ggplot2 `r packageVersion("ggplot2")`*
